---
title: "Reproducible Research Course Project 1"
author: "Heidi Peterson"
date: "10/7/2019"
output: 
  html_document: 
    keep_md: yes
---

variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken

Download & unzip dataset 
read dataset Process/transform the data 
(if necessary) into a format suitable for your analysis

```{r}
library(ggplot2)
library(lubridate)

temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
unzip(temp)
```

read dataset Process/transform the data 
(if necessary) into a format suitable for your analysis

```{r}

activityRR<- read.csv("activity.csv")
summary(activityRR)
activityRR$date<-ymd(activityRR$date)
summary(activityRR)
head(activityRR)
```

Histogram of the total number of steps taken each day

```{r}
stepseachDay <- tapply(activityRR$steps, activityRR$date, sum, na.rm=TRUE)

qplot(stepseachDay,
      geom="histogram",
      main = "Histogram of the total number of steps taken each day", 
      xlab = "Steps",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2))


hist(stepseachDay,col = "blue", main="Histogram of the total number of steps taken each day", xlab= "steps each day")
```

Mean and median number of steps taken each day

```{r}
summary(stepseachDay)
mean<-mean(stepseachDay)
median(stepseachDay)
```

Time series plot of the average number of steps taken had to make a new matrix as the old one was not labled columns

```{r}
totalstepsperday <- aggregate(steps ~ date, data = activityRR, FUN = sum, na.rm = TRUE)
head(totalstepsperday)
summary(totalstepsperday)

ggplot(totalstepsperday, aes(date, steps)) + geom_line() +
  xlab("") + ylab("Daily Steps")
```

The 5-minute interval that, on average, contains the maximum number of steps
The origional dataframe has 5 the steps measured in 5 minutes intervals per day.

```{r}

ggplot(activityRR, aes(x = interval , y = steps))+
  geom_line(color="blue", size=1) + 
  labs(title = "steps per 5 minute intervals", x = "5 min Interval", y = " Steps")

summary(activityRR)

```
We find out what the maximum amount of steps is by the function below which confirms with the summary chart above.

```{r}
max_step<- max(activityRR$steps, na.rm=T)
max_step
```

Then we place this in the datafram to find out which 5 minute interval contains that maximum number of steps and we find
```{r}
activityRR[max_step, ]$interval
```

Code to describe and show a strategy for imputing missing data
replacing missing values with the mean

```{r}
activityRR$steps[is.na(activityRR$steps)]<-mean(activityRR$steps, na.rm=TRUE)

```
Histogram of the total number of steps taken each day after missing values are imputed

```{r}
stepseachDay <- tapply(activityRR$steps, activityRR$date, sum, na.rm=TRUE)

qplot(stepseachDay,
      geom="histogram",
      main = "Histogram of the total number of steps taken each day", 
      xlab = "Steps",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2))


hist(stepseachDay,col = "blue", main="Histogram of the total number of steps taken each day", xlab= "steps each day")
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

find the days by the date
```{r}
activityRR$days<-weekdays(as.Date(activityRR$date))

```
Make a new column decerning between weekdays and weekends

```{r}
activityRR$wDay  <- ifelse(weekdays(activityRR$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")

```
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
ggplot(activityRR , aes(x = interval , y = steps)) +
  geom_line() +
  labs(title = "Steps by intervals", x = "interval", y = "steps") +
  facet_wrap(~wDay , ncol = 1, nrow=2)

```
All of the R code needed to reproduce the results (numbers, plots, etc.) in the report is above



